Выпускная квалификационная работа

### 2025-06-02 TeX/CDAM/en_cdam_KhromovGolyandina.tex
I L K problem

\subsection{Trajectory tensors}
Let $\tX = (x_1, x_2, \ldots, x_N)$ be a single-channel time series
of length $N$, $x_n \in \mathbb{C}$.
\begin{definition}
  The tensor embedding operator for a single-channel time series with
  window lengths $I$ and $J=L$ (then $K=N-I-L+2$) such that ${1< I,L < N},\, {I + L < N + 1}$
  is a mapping $\calT_{I,L}$ that transfers the series
  $\tX$ into the tensor $\calX \in \mathbb{C}^{I\times L \times K}$
  as follows: ${\mathcal{X}_{ilk}=x_{i+l+k-2}}$, where
  $i\in \overline{1:I},\, l \in\overline{1:L},\, k \in\overline{1:K}$.
\end{definition}

Let $\tX = (\tX^{(1)}, \tX^{(2)}, \ldots, \tX^{(P)})$ be a multi-channel
time series consisting of $P$ single-channel series, also called channels.
\begin{definition}
  The tensor embedding operator for a multi-channel time series with
  window length $L$ (that is, $I=P$, $J=L$, $K=N-L+1$) such that ${1< L < N}$ is a mapping $\calT_{L}$
  that translates the
  $P$-channel time-series $\tX$ into the tensor $\calX \in
  \mathbb{C}^{L\times K \times P}$ (${K = N - L + 1}$) as follows
  $\mathcal{X}_{lkp} =  x_{l+k-1}^{(p)}$, where ${l \in
  \overline{1:L}},\, {k \in \overline{1:K}},\, {p \in \overline{1:P}}$.
\end{definition}

@@

The following methods were compared for single channel time series
and signal extraction problem: SSA, HO-SSA, SSADstack, HO-SSADstack
with $R_3=r$ and HO-SSADstack with $R_3=1$. It has been shown that, in most cases, the SSA method significantly outperforms other methods in terms of accuracy. When the SSA method is less accurate, the difference is negligible and only occurs for a very narrow range of parameters. This minor disadvantage is therefore not a practical consideration. Of the Dstack methods, SSADstack and HO-SSADstack are the most accurate, with a small difference in accuracy when $R_3=r$.

For single-channel time series and frequency estimation problem, a
signal in the form of two sinusoids with close frequencies was
considered. The ESPRIT, HO-ESPRIT, HTLSDstack, HO-HTLSDstack with
$R_3=r$ and HO-HTLSDstack with $R_3=1$ methods were compared. It was found that ESPRIT performs more accurately at a low noise level. However, at a medium or high noise level, HO-ESPRIT with optimal parameter selection becomes more accurate. Furthermore, HO-HTLSDstack with $R_3=1$ outperforms all methods.

>>>>

\subsection{Trajectory tensors}
Let $\tX = (x_1, x_2, \ldots, x_N)$ be a single-channel time series
of length $N$, $x_n \in \mathbb{C}$.
\begin{definition}
  The tensor embedding operator for a single-channel time series with
  window lengths $I$ and $L$ (then $K=N-I-L+2$) such that ${1< I,L < N},\, {I + L < N + 1}$
  is a mapping $\calT_{I,L}$ that transfers the series
  $\tX$ into the tensor $\calX \in \mathbb{C}^{I\times L \times K}$
  as follows: ${\mathcal{X}_{ilk}=x_{i+l+k-2}}$, where
  $i\in \overline{1:I},\, l \in\overline{1:L},\, k \in\overline{1:K}$.
\end{definition}

Let $\tX = (\tX^{(1)}, \tX^{(2)}, \ldots, \tX^{(P)})$ be a multi-channel
time series consisting of $P$ single-channel series, also called channels.
\begin{definition}
  The tensor embedding operator for a multi-channel time series with
  window length $L$ (that is, $I=P$, $L$, $K=N-L+1$) such that ${1< L < N}$ is a mapping $\calT_{L}$
  that translates the
  $P$-channel time-series $\tX$ into the tensor $\calX \in
  \mathbb{C}^{L\times K \times P}$ (${K = N - L + 1}$) as follows
  $\mathcal{X}_{lkp} =  x_{l+k-1}^{(p)}$, where ${l \in
  \overline{1:L}},\, {k \in \overline{1:K}},\, {p \in \overline{1:P}}$.
\end{definition}

@@

The following methods were compared for single channel time series
and signal extraction problem: SSA, HO-SSA, SSADstack, HO-SSADstack
with $R_1=r$ and HO-SSADstack with $R_1=1$. It has been shown that, in most cases, the SSA method significantly outperforms other methods in terms of accuracy. When the SSA method is less accurate, the difference is negligible and only occurs for a very narrow range of parameters. This minor disadvantage is therefore not a practical consideration. Of the Dstack methods, SSADstack and HO-SSADstack are the most accurate, with a small difference in accuracy when $R_1=r$.

For single-channel time series and frequency estimation problem, a
signal in the form of two sinusoids with close frequencies was
considered. The ESPRIT, HO-ESPRIT, HTLSDstack, HO-HTLSDstack with
$R_1=r$ and HO-HTLSDstack with $R_1=1$ methods were compared. It was found that ESPRIT performs more accurately at a low noise level. However, at a medium or high noise level, HO-ESPRIT with optimal parameter selection becomes more accurate. Furthermore, HO-HTLSDstack with $R_1=1$ outperforms all methods.
